---
title: "W271 Lab3 - Adjusted Charges Regression"
author: "Steve Pelkey"
date: "December 6, 2016"
output: html_document
---
```{r}
#Initial data load and imports
library(dplyr)
library(tidyr)
library(ggplot2)
library(stargazer)
library(car)
library(lmtest)
library(ivpack)
library(effects)


setwd("c:/Users/Steve/Documents/mids/w_271/project/W271/Data/R/")
d <- read.delim("../../../W271_Lab3_Dataset_By_Hosp.txt")
```


##2013 Adjusted Charges

The primary independent variable is the adjusted charges of a provider relative to the charges of the other providers in its HRR in 2013. Adjusting charges based on the price and complexity of a service is a standard practice in the healthcare industry designed to make comparing hospitals easier. As demonstrated by the plot below, there is a strong positive relationship between what a provider charges and its competitors (r = .66).

```{r}
# clear linear relationship between average charges of provider and its competitors
ggplot(d, aes(FY2013_ADJUSTED_AVG_CHG, HRR_FY2013_ADJUSTED_AVG_CHG_EXCLUDING_HOSP)) +
  geom_point() + geom_smooth(method = 'lm') +
  labs(x = 'Average Provider Charges ($)', 
       y = 'Average Charges of Competitors($)', 
       title = 'Adjusted Average Adjusted Charges for Provider and Its HRR')

#correlation of .66
cor(d$FY2013_ADJUSTED_AVG_CHG, d$HRR_FY2013_ADJUSTED_AVG_CHG_EXCLUDING_HOSP, use = "pairwise.complete.obs")
```

Despite this relationship, there is still a lot variance in charges within an HRR. Provider charges ranged from -85 to 208% of their average HRR charge, with an average of -7.071% and a standard deviation of 36 percentage points. The distribution is mostly normal, but has a positive skew, which statistically deviates from normality (W =.96, p < .001) This will serve as a our primary independent variable as a proxy for local market competition.

```{r}
stargazer(d$FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX %>% data.frame(),
          type = "text", iqr = T, median = T,
          title = "Provider Adjusted Charges Relative to Competitors, 2013")

#asses distribution
#distribution kind of normal looking. slight positive skew
par(mfrow=c(1,2))
hist(d$FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX, breaks = 40,
     main = "Charges as Percent of Competitors")
qqPlot(d$FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX)
par(mfrow=c(1,1))
shapiro.test(d$FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX)
```

##Difference in Adjusted Charges 2013-14

For the most part, adjusted charges between 2013 and 2014 did not change too dramatically. On average, adjusted charges increased by $1,054 with a standard deviation of $2,548. Notably, there are a few outliers on either side of the distribution, which makes it leptokurtic This will serve as a our dependent variable.
```{r}
#look at relationship between 2013 and 14 charges

ggplot(d, aes(FY2013_ADJUSTED_AVG_CHG, FY2014_ADJUSTED_AVG_CHG)) + 
  geom_point() + geom_smooth(method = 'lm') +
  labs(x = "2013 Adjusted Average Charge", y = "2014 Adjusted Average Charge") +
  ggtitle("Provider Adjusted Average Charges, 2013 by 2014")

#.99 correlation
cor(d$FY2013_ADJUSTED_AVG_CHG, d$FY2014_ADJUSTED_AVG_CHG)

#summarise change variable
stargazer(d$DIFF_14_13_ADJUSTED_AVG_CHG %>% data.frame(),
          type = "text", iqr = T, median = T,
          title = "Charge in Provider Charges, 2013-2014")

#asses distribution
#distribution leptokurtic
par(mfrow=c(1,2))
hist(d$DIFF_14_13_ADJUSTED_AVG_CHG, breaks = 40,
     main = "Charge in Provider Charges, 2013-2014")
qqPlot(d$DIFF_14_13_ADJUSTED_AVG_CHG)
par(mfrow=c(1,1))

shapiro.test(d$DIFF_14_13_ADJUSTED_AVG_CHG)

```

##Bivariate EDA

In order to assess our hypothesis that local market competition is a primary driver of healthcare inflation, we examined the relationship between a provider's adjusted charges relative to its HRR in 2012 to how it set its charges for 2014 (in comparison to its 2013 charges). In a healthy market, we expect a negative relationship; those providers who were relatively costly should either increase their charges less dramatically or lower them to remain competitive in their local market. However, the scatterplot below reveals a weak positive trend (r = .06). Even though this relationship is statistically significant (p <.001) it isn't very practically significant. For every one percent increase in 2012 provider adjusted charges relative to its HRR, adjusted charges increased by just $5 from 2013 to 2014, which is just .2% of the standard deviation of the dependent variable. The tiny r2 indicates that much of the variance in 2014 adjusted charges has yet to be explained, so we added covariates to our model.

```{r}
#Weak positive relationship
ggplot(d, aes(FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX, DIFF_14_13_ADJUSTED_AVG_CHG)) + 
  geom_point() + geom_smooth(method = 'lm') +
  ggtitle("Increase in Provider Charge in 2014 by Relative 2013 Charge") +
  labs(x = "2013 Average Charge Relative to Competition (Percent)", y = "Difference in Charge, 2013 - 14")

#Tiny correlation
cor(d$FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX, d$DIFF_14_13_ADJUSTED_AVG_CHG, use = "pairwise.complete.obs")

# Statistically significant
adj_model0 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX)
stargazer(adj_model0, 
          type = "text",
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "Bivariate Model",
          covariate.labels = c("Average Relative Charge, 2013"))

# Not hugely practically significant
# coefficient is .24% of one sd of dependent variable
adj_model0$coefficients[2]/sd(d$DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T) * 100

hist(adj_model0$residuals, breaks = 50)

#diagnostic plots. heteroskedasticity might be a problem
par(mfrow = c(2,2))
plot(adj_model0)
par(mfrow = c(1,1))

#just significant
bptest(adj_model0)

#use robust SEs just to be safe
se_adj_model0 = robust.se(adj_model0)[,2]

stargazer(adj_model0, 
          type = "text",
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "Bivariate Model",
          covariate.labels = c("Average Relative Charge, 2013"),
          se = list(se_adj_model0 ))
```

##Building the adjusted charges model

###Estimating Competitiveness of Local Markets

In theory, the size of the HRR market relative to the number of providers should influence adjusted charges since a relatively crowded market should be more competitive. We used the average number of Part A beneficiaries per hospital in an HRR as a proxy for this phenomenon. Indeed, we observed a positive relationship in bivariate analysis that was preserved when added to the regression, and is statistically and practically significant. There is a $31.75 dollar increase for every 1,000 Part A beneficiaries per provider in an HRR. Thus, this will remain in our model.

```{r}
stargazer(d$BENEFICIARIES_PER_PROVIDER %>% data.frame(),
          type = "text", iqr = T, median = T,
          title = "Part A Beneficiaries per Provider")

#distribution
par(mfrow = c(1,2))
hist(d$BENEFICIARIES_PER_PROVIDER, breaks = 20)
qqPlot(d$BENEFICIARIES_PER_PROVIDER)
par(mfrow = c(1,1))

#slight positive relationship with dependent variable
ggplot(d, aes(BENEFICIARIES_PER_PROVIDER, DIFF_14_13_ADJUSTED_AVG_CHG)) + 
  geom_point() + geom_smooth(method = 'lm') +
  ggtitle("Beneficiaries per Provider by Difference in Adjusted Charges, 2013-14") +
  labs(x = "Beneficiaries per Provider", y = "Difference in Adjusted Average Charge, 2013-14")
cor(d$BENEFICIARIES_PER_PROVIDER, d$DIFF_14_13_ADJUSTED_AVG_CHG, use = "pairwise.complete.obs")

#add to model. Slight improvement
adj_model1 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER)

se_adj_model1 = robust.se(adj_model1)[,2]

stargazer(adj_model0, adj_model1,
          type = "text", 
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "With Beneficiaries per Provider",
          covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)"),
          se = list(se_adj_model0, se_adj_model1 ))
AIC(adj_model0,adj_model1)

```

###Bed Count Per Thousand Beneficiary

Bed count is a proxy for the size of the provider. While we expect it to be positively correlated with average number of Part A beneficiaries per provider (r = .16), a good part of the variation in this correlation represents the fact that there are major discrepancies in provider size within HRRs. This discrepancy is likely a factor in local market competition when providers are setting charges. After computing the size of a provider relative to its HRR and entering this variable into the model, we found that the AIC dropped and R2 increased considerably. Furthermore, adding in this variable greatly reduced the coefficient of our primary independent variable, which is no longer statistically significant.

```{r}
#summary stats
stargazer(d$BED_CNT_PER_THOUSAND_BENEFICIARIES %>% data.frame(),
          type = "text", iqr = T, median = T,
          title = "Bed Count per Thousand Part A Beneficiaries")

#distribution
#Poisson-esque
par(mfrow = c(1,2))
hist(d$BED_CNT_PER_THOUSAND_BENEFICIARIES, breaks = 20,
     main = "Bed Count per Thousand Part A Beneficiaries")
qqPlot(d$BED_CNT_PER_THOUSAND_BENEFICIARIES)
par(mfrow = c(1,1))

#slight positive relationship with dependent variable
ggplot(d, aes(BED_CNT_PER_THOUSAND_BENEFICIARIES, DIFF_14_13_ADJUSTED_AVG_CHG)) + 
  geom_point() + geom_smooth(method = 'lm') +
  ggtitle("Bed Count per Thousand Part A Beneficiaries by Difference in Adjusted Charges, 2013-14") +
  labs(x = "Bed Count per Thousand Part A Beneficiaries", y = "Difference in Adjusted Average Charge, 2013-14")
cor(d$BED_CNT_PER_THOUSAND_BENEFICIARIES, d$DIFF_14_13_ADJUSTED_AVG_CHG, use = "pairwise.complete.obs")

#add to model
adj_model2 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES)

se_adj_model2 = robust.se(adj_model2)[,2]

stargazer(adj_model1,adj_model2,
          type = "text", 
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "With Beds per Thousand Part A Beneficiaries",
          covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries"),
          se = list(se_adj_model1, se_adj_model2 ))
AIC(adj_model1,adj_model2)

```


###Urban vs Rural

It is well established that urban areas are more expensive than rural areas, and could thus have a large impact on provider charges. There are two variables in the dataset denoting rural/urban discrepancies: 1) percent of providers in an HRR that are urban, 2) the urban/rural status of an individual provider. Either or both (potential interaction) could have an impact on adjusted charges. 

On average, urban providers increased their charges at nearly twice the rate of rural providers between 2013 and 2014, and there is a small positive correlation between adjusted charges and the percent of HRR that is urban (r = .09). We tried out different combinations of these variables in our mode. Both were significant by themselves, only HRR urban percentage was significant when added together, and when an interaction effect was added, it was not significant. We selected the model with this lowest AIC (56893), which was just HRR urban percentage.

```{r}
#Summary stats and plots
#summary stats by rural/urban
stargazer(d %>% group_by(URBAN_RURAL_IND) %>%
  summarise(count = n(),
            avg = mean(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            median = median(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            stdv = sd(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            min = min(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            max = max(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T)) %>%
    data.frame(), type = "text", summary = F, 
  title = "Difference in Adjusted Average Charge, 2013-14 by Urban/Rural" )

#density plot. massive overlap between rural and urban
ggplot(d, aes(DIFF_14_13_ADJUSTED_AVG_CHG, fill = URBAN_RURAL_IND, colour = URBAN_RURAL_IND)) + 
  geom_density(alpha = .1) +
  ggtitle("Difference in Adjusted Average Charge, 2013-14 by Urban/Rural") +
  labs(x = "Difference in Adjusted Average Charge, 2013-14")

#Provider Status and Percent urban in HRR
#huge variability in the 100% urban section
#possible interaction
ggplot(data = d, aes(HRR_PERCENT_URBAN, DIFF_14_13_ADJUSTED_AVG_CHG, colour = factor(URBAN_RURAL_IND))) + 
  geom_point() + geom_smooth(method = 'lm') + 
  labs(colour = "Provider \nU/R Status", 
         x ="Percent of HRR that is Urban",
         y = "Difference in Adjusted Average Charge, 2013-14") +
  ggtitle("Difference in Adjusted Average Charge, 2013-14\n by Urban/Rural Status and Percent of HRR That is Urban")
cor(d$HRR_PERCENT_URBAN,d$DIFF_14_13_ADJUSTED_AVG_CHG)

#compare models
#provider urban/rural status
adj_model3 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  urban_dummy)

se_adj_model3 = robust.se(adj_model3)[,2]


#percent of HRR that is urban instead
adj_model4 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN)
se_adj_model4 = robust.se(adj_model4)[,2]

#both urban/rural variables
adj_model5 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  urban_dummy +
                  HRR_PERCENT_URBAN)
se_adj_model5 = robust.se(adj_model5)[,2]

#both urban/rural variables with interaction
adj_model6 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  urban_dummy +
                  HRR_PERCENT_URBAN +
                  urban_dummy * HRR_PERCENT_URBAN)
se_adj_model6 = robust.se(adj_model6)[,2]

stargazer(adj_model3, adj_model4,adj_model5,adj_model6,
          type = "text",
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "With Urban/Rural Variables",
          covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries",
                               "Urban Provider",
                               "Percent of HRR that is Urban",
                               "Urban by Percent Urban Interaction"),
          se = list( se_adj_model3, se_adj_model4,se_adj_model5,se_adj_model6  ))
AIC(adj_model3, adj_model4,adj_model5,adj_model6 )


```

###Residency

Same deal as urban just with residency. write later. Neither are significant by themselves, but when added together with an interaction, they are. Adding all three significantly improved the model (f = 2.95, p < .05).

```{r}
#summary stats by residency
stargazer(d %>% group_by(RESIDENCY_PROGRAM_IND) %>%
  summarise(count = n(),
            avg = mean(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            median = median(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            stdv = sd(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            min = min(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            max = max(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T)) %>%
    data.frame(), type = "text", summary = F, 
  title = "Providers With/Out Residency Programs")

#density plot
ggplot(d, aes(DIFF_14_13_ADJUSTED_AVG_CHG, fill = RESIDENCY_PROGRAM_IND, colour = RESIDENCY_PROGRAM_IND)) + 
  geom_density(alpha = .1) + 
  labs(x = "Difference in Adjusted Average Charge, 2013-14") +
  ggtitle("Difference in Adjusted Average Charge, 2013-14\n by Provider Residency Status")

#same trend as urban/rural, though interaction effect looks potentially significant
#providers who are the only 
ggplot(data = d, aes(HRR_PERCENT_RESIDENT, DIFF_14_13_ADJUSTED_AVG_CHG, colour = factor(RESIDENCY_PROGRAM_DUMMY))) + 
  geom_point() + geom_smooth(method = 'lm') + 
  labs(colour = "Residency", x = "Percent of HRR that Has Residency Program", y = "Difference in Adjusted Average Charge, 2013-14") +
  ggtitle("Difference in Adjusted Average Charge, 2013-14\n by Residency Status and Percent of HRR That has Resident Program")
cor(d$HRR_PERCENT_RESIDENT,d$DIFF_14_13_ADJUSTED_AVG_CHG)

#compare models
#just provider residency status
adj_model7 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  RESIDENCY_PROGRAM_DUMMY)
se_adj_model7 = robust.se(adj_model7)[,2]

#just percent residency
adj_model8 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  HRR_PERCENT_RESIDENT )
se_adj_model8 = robust.se(adj_model8)[,2]

#both provider residency status and percent residency
adj_model9 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  HRR_PERCENT_RESIDENT +
                  RESIDENCY_PROGRAM_DUMMY)
se_adj_model9 = robust.se(adj_model9)[,2]

#add interaction
adj_model10 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  HRR_PERCENT_RESIDENT +
                  RESIDENCY_PROGRAM_DUMMY +
                  HRR_PERCENT_RESIDENT * RESIDENCY_PROGRAM_DUMMY)
se_adj_model10 = robust.se(adj_model10)[,2]

stargazer(adj_model7, adj_model8,adj_model9,adj_model10,
          type = "text", 
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "With Urban/Rural Variables",
          covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries",
                               "Percent of HRR that is Urban",
                               "Has Residency Program",
                               "Provider Residency by Percent of HRR w Residency",
                               "Percent of HRR with Residency Programs"),
          se = list(se_adj_model7, se_adj_model8,se_adj_model9, se_adj_model10))
AIC(adj_model4, adj_model7, adj_model8,adj_model9,adj_model10 )

#new model is significantly better
linearHypothesis(adj_model10, c("HRR_PERCENT_RESIDENT=0", "RESIDENCY_PROGRAM_DUMMY =0","HRR_PERCENT_RESIDENT:RESIDENCY_PROGRAM_DUMMY=0"))

```

###Discharges per bed count/Health of Provider

Discharges per bed is a proxy for the health of provider (pun intended). More discharges per bed means that a hospital is doing better business relative to its capacity. We used two variables for discharges per bed count. 1) 2013 discharges per bed gauges the health of provider against other providers, 2) the difference in discharges between 2013 and 2012 gauges the health of a provider relative to itself. These are jointly significant.

```{r}
#2013 discharges
#potential outlier situation. see what it looks like in final model
ggplot(d, aes(FY2013_DISCHARGES_PER_BED, DIFF_14_13_ADJUSTED_AVG_CHG)) + 
  geom_point() + geom_smooth(method = 'lm') +
  labs(x = "2013 Discharges per Bed", y = "Difference in Adjusted Average Charge, 2013-14") +
  ggtitle("Difference in Adjusted Average Charge, 2013-14\n by Number of Discharges per Bed, 2013")

#change in discharges, 2012-13 
#potential outlier situation. see what it looks like in final model
ggplot(d, aes(DIFF_13_12_DISCHARGES_PER_BED, DIFF_14_13_ADJUSTED_AVG_CHG)) + 
  geom_point() + geom_smooth(method = 'lm') +
  labs(x = "Difference in Discharges per Bed,2012-13", y = "Difference in Adjusted Average Charge, 2013-14") +
  ggtitle("Difference in Adjusted Average Charge, 2013-14\n by Difference in Discharges per Bed, 2012-13")

adj_model11 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  HRR_PERCENT_RESIDENT +
                  RESIDENCY_PROGRAM_DUMMY +
                  HRR_PERCENT_RESIDENT * RESIDENCY_PROGRAM_DUMMY +
                  FY2013_DISCHARGES_PER_BED +
                  DIFF_13_12_DISCHARGES_PER_BED)
se_adj_model11 = robust.se(adj_model11)[,2]

stargazer(adj_model10, adj_model11,
          type = "text", 
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "With Discharges per Bed",
          covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries",
                               "Percent of HRR that is Urban",
                               "Percent of HRR with Residency Programs",
                               "Has Residency Program",
                               "Discharges per Bed, 2013",
                               "Discharges per bed, Change from 2012-13",
                               "Provider Residency by Percent of HRR w Residency"),
          se = list(se_adj_model10, se_adj_model11 ))
AIC(adj_model10, adj_model11)

#jointly significant
linearHypothesis(adj_model11, c("FY2013_DISCHARGES_PER_BED=0", "DIFF_13_12_DISCHARGES_PER_BED =0"))

```

###Change in Ownership

Did provider go through a change in ownership. Very significant, statistically and practically so. 
```{r}
#summary stats by change in ownership status
stargazer(d %>% group_by(CHANGE_OF_OWNERSHIP_IN_PERIOD) %>%
  summarise(count = n(),
            avg = mean(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            median = median(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            stdv = sd(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            min = min(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            max = max(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T)) %>%
    data.frame(), type = "text", 
  title = "Change in Ownership?",summary = F)

adj_model12 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  HRR_PERCENT_RESIDENT +
                  RESIDENCY_PROGRAM_DUMMY +
                  HRR_PERCENT_RESIDENT * RESIDENCY_PROGRAM_DUMMY +
                  FY2013_DISCHARGES_PER_BED +
                  DIFF_13_12_DISCHARGES_PER_BED +
                  CHANGE_OF_OWNERSHIP_IN_PERIOD)
se_adj_model12 = robust.se(adj_model12)[,2]

stargazer(adj_model11, adj_model12,
          type = "text", 
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "With Change in Ownership",
          covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries",
                               "Percent of HRR that is Urban",
                               "Percent of HRR with Residency Programs",
                               "Has Residency Program",
                               "Discharges per Bed, 2013",
                               "Discharges per bed, Change from 2012-13",
                               "Underwent Change in Management",
                               "Provider Residency by Percent of HRR w Residency"),
          se = list(se_adj_model11, se_adj_model12 ))
AIC(adj_model11, adj_model12)
```

Is a provider a private for profit? 
```{r}
#21% of providers are private for profits
mean(d$OWNERSHIP_FOR_PROFIT)

#summary stats by change in ownership status
stargazer(d %>% group_by(OWNERSHIP_FOR_PROFIT) %>%
  summarise(count = n(),
            avg = mean(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            median = median(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            stdv = sd(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            min = min(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T),
            max = max(DIFF_14_13_ADJUSTED_AVG_CHG, na.rm = T)) %>%
    data.frame(), type = "text", 
  title = "Private, For Profit Provider",summary = F)

#add ownership type
adj_model13 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  HRR_PERCENT_RESIDENT +
                  RESIDENCY_PROGRAM_DUMMY +
                  HRR_PERCENT_RESIDENT * RESIDENCY_PROGRAM_DUMMY +
                  FY2013_DISCHARGES_PER_BED +
                  DIFF_13_12_DISCHARGES_PER_BED +
                  CHANGE_OF_OWNERSHIP_IN_PERIOD +
                  OWNERSHIP_FOR_PROFIT )
se_adj_model13 = robust.se(adj_model13)[,2]

#sucked all significance from primary independent variable
stargazer(adj_model12, adj_model13,
          type = "text", 
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "With Ownership Type",
                    covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries",
                               "Percent of HRR that is Urban",
                               "Percent of HRR with Residency Programs",
                               "Has Residency Program",
                               "Discharges per Bed, 2013",
                               "Discharges per bed, Change from 2012-13",
                               "Underwent Change in Management",
                               "Is Private, For Profit",
                               "Provider Residency by Percent of HRR w Residency"),
          se = list(se_adj_model12, se_adj_model13 ))
AIC(adj_model12, adj_model13)

#interaction effect with primary independent variable?
ggplot(data = d, 
       aes(FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX, DIFF_14_13_ADJUSTED_AVG_CHG, colour = factor(OWNERSHIP_FOR_PROFIT))) + 
  geom_point() + geom_smooth(method = 'lm') + 
  labs(colour = "Private\nfor Profit?", x = "Average Relative Charge, 2013", y = "Difference in Adjusted Average Charge, 2013-14") +
  ggtitle("Average Relative Charge, 2013 by Difference in Adjusted Average Charge, 2013-14 \n by Ownership Type")

#add interaction effect
adj_model14 = lm(data = d, DIFF_14_13_ADJUSTED_AVG_CHG ~ FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  HRR_PERCENT_RESIDENT +
                  RESIDENCY_PROGRAM_DUMMY +
                  HRR_PERCENT_RESIDENT * RESIDENCY_PROGRAM_DUMMY +
                  FY2013_DISCHARGES_PER_BED +
                  DIFF_13_12_DISCHARGES_PER_BED +
                  CHANGE_OF_OWNERSHIP_IN_PERIOD +
                  OWNERSHIP_FOR_PROFIT +
                  OWNERSHIP_FOR_PROFIT * FY2013_ADJUSTED_AVG_CHG_RELATIVE_EX )
se_adj_model14 = robust.se(adj_model14)[,2]

stargazer(adj_model12, adj_model13, adj_model14,
          type = "text", 
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "With Ownership Type, Relative Charge Interaction",
                    covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries",
                               "Percent of HRR that is Urban",
                               "Percent of HRR with Residency Programs",
                               "Has Residency Program",
                               "Discharges per Bed, 2013",
                               "Discharges per bed, Change from 2012-13",
                               "Underwent Change in Management",
                               "Is Private, For Profit",
                               "Provider Residency by Percent of HRR w Residency",
                               "For Profit by 2013 Relative Charge"),
          se = list(se_adj_model12, se_adj_model13, se_adj_model14 ))
AIC(adj_model12, adj_model13, adj_model14)

```

Final model for nonsplit sample.

```{r}
overall_model = adj_model14
se_overall_model = se_adj_model14

stargazer(overall_model,
          type = "text", 
          dep.var.labels = "Difference in Adjusted Average Charge, 2013-14",
          title = "Final Overall Model",
                    covariate.labels = c("Average Relative Charge, 2013", 
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries",
                               "Percent of HRR that is Urban",
                               "Percent of HRR with Residency Programs",
                               "Has Residency Program",
                               "Discharges per Bed, 2013",
                               "Discharges per bed, Change from 2012-13",
                               "Underwent Change in Management",
                               "Is Private, For Profit",
                               "Provider Residency by Percent of HRR w Residency",
                               "For Profit by 2013 Relative Charge"),
          se = list(se_overall_model))
stargazer(adj_model14,
          type = "text", se = list(se_adj_model13 ))
```

Adjusted vs non adjusted dollars. Most important effect is that the primary independent variable remained significant with unadjusted dollars.

```{r}

#use unadjusted dollars
non_adj_model = lm(data = d, DIFF_14_13_AVG_CHG ~ FY2013_AVG_CHG_RELATIVE_EX +
                  BENEFICIARIES_PER_PROVIDER + 
                  BED_CNT_PER_THOUSAND_BENEFICIARIES +
                  HRR_PERCENT_URBAN +
                  HRR_PERCENT_RESIDENT +
                  RESIDENCY_PROGRAM_DUMMY +
                  HRR_PERCENT_RESIDENT * RESIDENCY_PROGRAM_DUMMY +
                  FY2013_DISCHARGES_PER_BED +
                  DIFF_13_12_DISCHARGES_PER_BED +
                  CHANGE_OF_OWNERSHIP_IN_PERIOD +
                  OWNERSHIP_FOR_PROFIT +
                  OWNERSHIP_FOR_PROFIT * FY2013_AVG_CHG_RELATIVE_EX )
se_adj_modelnonad = robust.se(non_adj_model)[,2]

stargazer(overall_model,non_adj_model,
          type = "text", 
          dep.var.labels = c("Adjusted", "Non Adjusted"),
          title = "Adjusted vs Unadjusted Dollars",
                    covariate.labels = c("Average Relative Charge (Adjusted), 2013", 
                               "Average Relative Charge (Unadjusted), 2013",
                               "Avg Number of Beneficiaries per Provider (thousands)",
                               "Beds per Thousand Beneficiaries",
                               "Percent of HRR that is Urban",
                               "Percent of HRR with Residency Programs",
                               "Has Residency Program",
                               "Discharges per Bed, 2013",
                               "Discharges per bed, Change from 2012-13",
                               "Underwent Change in Management",
                               "Is Private, For Profit",
                               "Provider Residency by Percent of HRR w Residency",
                               "For Profit by 2013 Relative Charge (Adjusted)",
                               "For Profit by 2013 Relative Charge (Unadjusted)"),
          se = list(se_overall_model,se_adj_modelnonad ))

```
